Lightweight Real-Time Feature Monitoring

### 2020-07-19 paper.tex
Experiments, images scaled to two-column format

%
% The abstract is a short summary of the work to be presented in the article.
\begin{abstract}
Many real-time stream monitoring systems are static once deployed in a production environment. Over time, due to data pattern shifts, the initially deployed static system's performance gradually deteriorates. Accurate and timely detection of data pattern deviations allow for immediate measures to be taken. Thus, the problem at hand is to determine when to reconfigure the system, \textit{e.g.}, a Machine Learning model, based on an analysis of the drifts in the stream of data. In this paper, we present a method that makes use of constant time and memory streaming aggregations and distribution divergence functions to alert deviations in data patterns in real-time. We generate alerts based on a user-defined probability threshold, for each of an event's numerical fields, denominated features in our context. We evaluate our method through a series of tests, some with synthetic data sets and some with real data. We further split the tests into single and multi-feature analysis. Our method accurately detected the introduced anomalies in the experiments using synthetic datasets while maintaining high throughput, for single and multi-feature analysis. However, experiments with real data were not as accurate. Despite not knowing the specific reasons for that, we detail further investigation to future work and formulate a set of hypotheses that might explain the latter results and hence are worthy of pursuing next.
\end{abstract}

@@

\section{Introduction}
In the past decade, applications have become increasingly data-driven, placing data at the center of application design. For different use cases, the data is processed in different ways for different purposes. For instance, e-commerce platforms such as Alibaba need to process a large number of daily transactions while ensuring that sales run smoothly and that products are delivered to customers' homes. Streaming services, like YouTube and Netflix, try to guarantee that media content reaches up to millions of users simultaneously. Social networks are responsible for generating large volumes of data. Twitter, for example, generates more than 500 million \textit{tweets} (posts in Twitter) per day. \textit{Tweets} may contain text, media content or both. Cybersecurity applications are yet another example of time-sensitive, data-driven applications. The entities that work in these use cases need to monitor user accesses and their actions in the network, where timely detection of intruders is critical to prevent them from tampering with the underlying system.

@@

\section{Method}
\label{sec:Method}
A data stream is a continuous collection of events, each timestamped and with multiple fields. In this paper, we refer to the event's fields as features. Our goal is to detect data pattern deviations for each feature in a streaming fashion. To that end, we devise a method that works with two windows and comprises two phases, a batch analysis phase and a stream analysis one. The batch analysis is performed on a static in time window which we call the reference window. The stream analysis is performed over a exponential tail which emulates a sliding window that groups data stream events. We aggregate the contents of the reference window with an exact histogram and the sliding window with an approximate histogram aggregation based on Exponential Moving Averages (EMAs). We then apply a statistical distribution divergence function to both histogram aggregations to obtain a divergence value. In the batch phase, we build the distribution of these values so we can define a percentile threshold for the newly measured divergence values in streaming. We raise alerts for features above certain percentiles, \textit{e.g.}, the 99th percentile.

@@

\subsection{Batch Analysis Phase} \label{sec:batch-phase}
The main goals of the batch phase are to build our reference aggregation and the distribution of divergence values, for each feature. Streaming divergence values will be compared to this distribution of values later to find out how probable a value is to occur (if it is very low, an alert is raised).

@@

\subsubsection{Building the Distribution of Divergence Values} \label{sec:sampling-batch}
Secondly, we aim to build the distribution of expected distance or divergence values, to later on threshold the observed distance values during the streaming phase. To that end, we make $S$ samples of transactions, each with the same tuple-based size of the sliding window we emulate in streaming. Each sample is a contiguous block of transactions, thus preserving the order of transactions and the time-dependency property of a time-series. With this sampling we aim to build a distribution of divergence values and reliably estimate the upper tail of that distribution. In \cite{SAMM}, the authors show how many samples must be made to have at least one value in the upper tail of the distribution. Defining $\alpha$ as our threshold (and $\frac{\alpha}{n_{features}}$ as our threshold corrected by the Holm-Bonferroni multiple test correction \cite{HolmBonferroni}, detailed in Section \ref{sec:multi-test}), $\gamma$ is defined as the probability of \textbf{not} having at least one point in the $\frac{\alpha}{n_{features}}$ upper tail of the distribution and the minimum number of samples $S$ to make is given by:
\begin{equation*}
    \label{eq:optimal-n-samples}
    \textit{S} = \ceil[\bigg]{\frac{\log\gamma}{log(1 - \frac{\alpha}{n_{features}})}}
\end{equation*}

@@

For each feature, we have now \textit{S} histograms, one for each sample. We compute the divergence between each of the \textit{S} histograms and the reference histogram. Figure \ref{fig:compute-sample-distances} illustrates this process as we apply our divergence or distance function for each sample's EMA-based histogram and the reference histogram, obtaining \textit{S} divergence measurements.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.4]{figures/compute-sample-distances.png}
      \caption[Compute sample's distance values]{Compute divergence values between reference and samples}
      \label{fig:compute-sample-distances}
    \end{center}
\end{figure}
These $S$ measurements are divergence measurements between random samples of data and the reference period. Hence, we claim we now know the distribution of expected distance values. Given a new distance value measured between another random window and the reference window, we can compute its probability and produce alerts if the probability is below a certain threshold. Note that the histograms for each sample are EMA-based histograms just like the streaming histogram. This ensures a certain degree of fidelity in the test we make in streaming because we measure the divergence between a random sample of data (our streaming histogram or any of the sample's histogram) and the reference one.

\subsubsection{Burn-In Period at System Initialization}
When we boot up our system in the streaming phase we initially have empty histogram aggregations. To avoid this burn-in period, we propose saving the last sample's histogram for each feature so that the streaming EMA-based histogram aggregation is initialized with it.

@@

\subsection{Streaming Phase} \label{sec:stream-phase}
In the streaming phase, the system will process event by event and perform a change detection test periodically. For each feature, we initialize the approximate histogram with the last sample's data, from the batch phase. Then, for each incoming event and each of the event's features, we update our streaming approximate histogram for that feature by adding the new value and then "cutting a slice" from the top of the histogram bars to simulate eviction, as described in Section \ref{sec:ema-hist}. This is all we do, regarding event processing.

We periodically perform a multi-feature change detection test. The test is done by first computing the divergence (here denoted as $div_x$) between the reference histogram for a feature \textit{x} and the current streaming EMA histogram for $x$. This value is tested against the known distribution of distance values for feature \textit{x}. We test to see if the percentile of the computed distance relative to the distribution of divergence is above a certain percentile. The probability of a divergence value $div_x$ is:
\[pvalue(div_x) = 1 - percentile(div_x)\]
For example, if we define alerts to have 1\% or less probability to happen, it means the system will raise an alert for distance values above the 99th-percentile or below a probability of \textit{p-value} = 0.01. We alert a feature \textit{x} as divergent at a given timestamp if for a given probability threshold $\alpha$:
\[ percentile(div_x) > 1 - \alpha \]
or:
\begin{equation}
    \label{eq:alert-test}
    pvalue(div_x) < \alpha
\end{equation}

\subsubsection{Multiple Test Correction} \label{sec:multi-test}
Note that we periodically apply this test (Equation \ref{eq:alert-test}) for each feature, effectively repeating the analysis. When considering several hypotheses, the problem of multiplicity arises \cite{MultiTestProblem-Pubmed, MultiTestProblem-Dickhaus2014, MultiTestProblem-Miller1966, multiple-test-correction-geoffrey}: the more hypotheses are checked, the higher the probability of obtaining false positives. In the literature, many multiple test correction methods are proposed, but we opted to use the Holm-Bonferroni \cite{HolmBonferroni} multiple test correction because it is a simple and easy to compute correction that can reduce the number of false negatives while having as target to control the number of false positives. Furthermore, it does not require to assume any independence between the tested hypothesis (as other methods do) which would almost surely be a wrong assumption for realistic datasets. In statistics, the Family-Wise Error Rate (FWER) \cite{MultivariateMT, MultitestTamhane2018AdvancesIP} is the probability of having one or more false positives when performing multiple hypotheses tests, \textit{i.e.}, considering one hypothesis true when it is not. The Holm-Bonferroni method is one of many approaches for controlling the Family-Wise Error Rate \textit{(FWER)} by adjusting the rejection criteria for each hypothesis. The Holm-Bonferroni correction takes Equation \ref{eq:alert-test} and corrects the right-side threshold $\alpha$ for each feature. This correction is done by dividing the threshold $\alpha$ by a correction factor. First, we compute the \textit{p-value} of each feature. We then order our features by \textit{p-value}, in ascending order. Next, starting from the first feature, the one with the smallest \textit{p-value}, and proceeding until the end of the list of features, we check if:
\begin{equation}
    pvalue(val_x) < \alpha / (m - k)
    \label{eq:corrected-pvalue}
\end{equation}
with $\alpha$ as the probability to reject the null hypothesis, \textit{m} as the number of hypotheses to test (in our case the number of features) and \textit{k} as the 0-based index of the current feature. We stop at the first feature that fails this test, and every feature processed up until this one is considered to be a divergent feature and an alert per feature is logged with corresponding probability value.

@@

\subsection{Single-Feature Analysis}
As an initial step, we started by validating our system using single-feature datasets before moving to multi-feature scenarios to facilitate system validation, considering the complexities of multi-feature datasets and the problem of multiple test correction.

\subsubsection{Varying Sample Sizes}
We began with single-feature scenarios and by experimenting with different tuple-based half-lives $n_{1/2}$. The half-life $n_{1/2}$ parameter in turn controls the size of the samples we make in the batch phase and the streaming EMA tail size. We set a half-life $n_{1/2}$ value and then multiply it by a constant factor of 4 to get the size of the samples to make. We process $4n_{1/2}$ events, meaning any further data we processed would make our oldest events have less than 6\% weight, following Algorithm \ref{pseudo:ema-hist}'s successively applied suppression factor. Processing $4n_{1/2}$ events means the oldest event's contribution is $({2^{- \frac{1}{n_{1/2}}}})^{4n_{1/2}} = {2^{- \frac{4n_{1/2}}{n_{1/2}}}} = 2^{-4} \approx 0.06 = 6\%$.

@@

From a functional point of view, were we able to detect the abrupt change introduced in the target dataset T1? Recall that from the 2.5 millionth event forward we change the distribution from a gaussian to an uniform one (Figure \ref{fig:timeseries-t1}). Figure \ref{fig:JSD-signal-01} shows the distance (JSD) values computed during streaming at each event of the 5 million events from dataset T1, using the outputs from the batch phase of experiment 01. We clearly see a change in JSD values, and we move from values close to 0 to values close to 1 (or even 1). Figure \ref{fig:JSD-signal-zoom-01} shows the same signal but for the first half of the streaming period (essentially a zoom-in). We see our JSD values before the change were very low (magnitudes of $10^{-2}$).
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-625.pdf}
      \caption{Whole time-series JSD signal and threshold}
      \label{fig:JSD-signal-01}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-zoom-625.pdf}
      \caption{Zoom-in on JSD signal and threshold}
     \label{fig:JSD-signal-zoom-01}
    \end{center}
\end{figure}
However, the change in JSD value is very abrupt as well. Almost instantly we go from close to 0 values to 1. We also had some false-positives, \textit{i.e.}, some JSD values above the threshold. This indicates to us that the window used is too small and sensitive to small changes, or that we can choose a more strict threshold.

In experiment 02, we use larger sampling windows. Remember our target sliding window in streaming is of the same size as the sampling windows. In the second experiment, we increase the size of the sampling windows by 100 times. Figure \ref{fig:JSD-signal-02} shows the JSD signal during streaming for dataset T1, using the outputs from the batch phase of experiment 02. Again, we see a change in signal from the middle of the streaming period forward. Figure \ref{fig:JSD-signal-zoom-02} shows the same signal for the first half of the streaming period.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-62500.pdf}
      \caption{Whole time-series JSD signal and threshold}
      \label{fig:JSD-signal-02}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-zoom-62500.pdf}
      \caption{Zoom-in on JSD signal and threshold}
     \label{fig:JSD-signal-zoom-02}
    \end{center}
\end{figure}
Notice that this time the growth of our signal is slower. Notice also that we do not have false positives in the first half of the signal. This is because we use a larger window, which means each event has a smaller contribution to the aggregation state. Hence, outliers will have less impact on the aggregation and corresponding signal value.

In experiment 03, we increase the sampling and target windows size once again, this time four times larger than in experiment 02. We would expect an even smoother growth of our signal and, again, no false positives. And indeed, judging by the corresponding signal computation and zoom-in, Figures \ref{fig:JSD-signal-03} and \ref{fig:JSD-signal-zoom-03}, respectively, the signal looks smoother and we see no false positives.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-250000.pdf}
      \caption{Whole time-series JSD signal and threshold}
      \label{fig:JSD-signal-03}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-zoom-250000.pdf}
      \caption{Zoom-in on JSD signal and threshold}
     \label{fig:JSD-signal-zoom-03}
    \end{center}
\end{figure}

\subsubsection{Multiple Distribution Changes}
In this experiment, we wanted to see how our signal behaved when there were multiple change points, where each gaussian distribution was a scaled down version of the previous one. Figure \ref{fig:timeseries-r2} shows the time-series for the reference dataset R2 used in this experiment. It contained 5 million events with a single feature following a gaussian distribution with mean $\mu=100$ and standard deviation $\sigma=50$. The time-series for target dataset T2 is shown in Figure \ref{fig:timeseries-t2}. Notice that we change distributions at 2, 3 and 4 millionth events. Notice also that each subsequent distribution produces values whose domain is a subset of the previous distribution.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/timeseries-r2.pdf}
      \caption{Time-series for dataset R2}
      \label{fig:timeseries-r2}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/timeseries-t2.pdf}
      \caption{Time-series for dataset T2}
     \label{fig:timeseries-t2}
    \end{center}
\end{figure}

The target dataset T2 begins with the same distribution as the reference dataset R2, a gaussian distribution with mean $\mu=100$ and standard deviation $\sigma=50$. At the 2 millionth event mark, we use another Gaussian distribution, this time with mean $\mu=100$ and standard deviation $\sigma=30$. At the 3 millionth event, we change the standard deviation to be $\sigma=10$. From the 4 millionth event to the end, we use a standard deviation of $\sigma=5$.

In this experiment, we used a half-life $n_{1/2}=62500$ and corresponding sample size of 250,000 events, like in the previous section's experiment 02. Figure \ref{fig:JSD-signal-test02} shows the computed JSD signal for the target dataset and Figure \ref{fig:JSD-signal-zoom-test02} zooms in on the first 2 millionth events, before any distribution is changed. We observe that once again we have no false positives in the first 2 million events, which is ideal. 
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-test02.pdf}
      \caption{Whole time-series JSD signal and threshold}
      \label{fig:JSD-signal-test02}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-zoom-test02.pdf}
      \caption{Zoom-in on first 2 million events}
     \label{fig:JSD-signal-zoom-test02}
    \end{center}
\end{figure}
Furthermore, we observe that the computed JSD signal varies accordingly with the changes in distribution. Despite us changing the distributions to subsets of the previous ones to try and deceive our method, our signal accurately represents these changes. In Figure \ref{fig:JSD-signal-test02}, we see several growth points that correspond to the points where the generating gaussian distribution parameters changed. Hence, we are able to detect changes in distribution parameter changes and quantify how big the change is (the closer to 1 our signal is).

\subsubsection{Multiple Distribution Changes, each a Superset of the Previous One}
In this experiment, we wanted to do the reverse of the previous one. In this one, there are multiple change points but each new distribution spans a domain that is a superset of the previous one. Figure \ref{fig:timeseries-r3} shows the reference period used, with 5 million events and one feature following a gaussian distribution of mean $\mu=100$ and standard deviation $\sigma=10$. Figure \ref{fig:timeseries-t3} shows the target period used, also with 5 million events and one feature. Like in the previous experiment, we change distributions at the 2, 3 and 4 millionth event mark. On the other hand, unlike the previous experiment, we now make subsequent distributions cover the previous ones. We change the reference gaussian used with mean $\mu=100$ and standard deviation $\sigma=10$ at the 2, 3 and 4 millionth mark, to use standard deviations $\sigma=30$, $\sigma=40$ and $\sigma=50$, respectively.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/timeseries-r3.pdf}
      \caption{Time-series for dataset R3}
      \label{fig:timeseries-r3}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/timeseries-t3.pdf}
      \caption{Time-series for dataset T3}
     \label{fig:timeseries-t3}
    \end{center}
\end{figure}

>>>>

%
% The abstract is a short summary of the work to be presented in the article.
\begin{abstract}
Many real-time stream monitoring systems are static once deployed in a production environment. Over time, due to data pattern shifts, the initially deployed static system's performance gradually deteriorates. Accurate and timely detection of data pattern deviations allow for immediate measures to be taken. Thus, the problem at hand is to determine when to reconfigure the system, \textit{e.g.}, a Machine Learning model, based on an analysis of the drifts in the stream of data. In this paper, we present a method that makes use of constant time and memory streaming aggregations and distribution divergence functions to alert deviations in data patterns in real-time. We generate alerts based on a user-defined probability threshold, for each of an event's numerical fields, denominated features in our context. We evaluate our method through a series of tests, some with synthetic datasets and some with real data. We further split the tests into single and multi-feature analysis. Our method accurately detected the introduced anomalies in the experiments using synthetic datasets while maintaining high throughput, for single and multi-feature analysis. However, experiments with real data were not as accurate. Despite not knowing the specific reasons for that, we detail further investigation to future work and formulate a set of hypotheses that might explain the latter results and hence are worthy of pursuing next.
\end{abstract}

@@

\section{Introduction}
In the past decade, applications have become increasingly data-driven, placing data at the center of application design. For different use cases, data is processed in different ways for different purposes. For instance, e-commerce platforms such as Alibaba need to process a large number of daily transactions while ensuring that sales run smoothly and that products are delivered to customers' homes. Live streaming services, like YouTube and Netflix, try to guarantee that media content reaches up to millions of users simultaneously. Social networks are responsible for generating large volumes of data: Twitter, for example, generates more than 500 million \textit{tweets} (posts in Twitter) per day. \textit{Tweets} may contain text, media content or both. Cybersecurity applications are yet another example of time-sensitive, data-driven applications. The entities that work in these use cases need to monitor user accesses and their actions in the network, where timely detection of intruders is critical to prevent them from tampering with the underlying system.

@@

\section{Method}
\label{sec:Method}
A data stream is a continuous collection of events, each timestamped and with multiple fields. In this paper, we refer to the event's fields as features. Our goal is to detect data pattern deviations for each feature in a streaming fashion. To that end, we devise a method that works with two windows and comprises two phases, a batch analysis phase and a stream analysis one. The batch analysis is performed on a static in time window which we call the reference window. The stream analysis is performed over an exponential tail aggregation which emulates a sliding window that groups data stream events. We aggregate the contents of the reference window with an exact histogram and the sliding window with an approximate histogram aggregation based on Exponential Moving Averages (EMAs). We then apply a statistical distribution divergence function to both histogram aggregations to obtain a divergence value. In the batch phase, we build the distribution of these values so we can define a percentile threshold for the newly measured divergence values in streaming. We raise alerts for features above certain percentiles, \textit{e.g.}, the 99th percentile.

@@

\subsection{Batch Analysis Phase} \label{sec:batch-phase}
The main goals of the batch phase are to build the reference histogram and the distribution of divergence values, for each feature. Streaming divergence values will be compared to this distribution of values later to find out how probable a value is to occur (if it is very low, an alert is raised).

@@

\subsubsection{Building the Distribution of Divergence Values} \label{sec:sampling-batch}
Secondly, we aim to build the distribution of expected distance or divergence values, to later on threshold the observed distance values during the streaming phase. To that end, we make $S$ samples of transactions of $K$ tuples ($S$ and $K$ detailed later). Each sample is a contiguous block of transactions, thus preserving the order of transactions and the time-dependency property of a time-series. With this sampling we aim to build a distribution of divergence values and reliably estimate the upper tail of that distribution. In \cite{SAMM}, the authors show how many samples must be made to have at least one value in the upper tail of the distribution. Defining $\alpha$ as our threshold (and $\frac{\alpha}{n_{features}}$ as our threshold corrected by the Holm-Bonferroni multiple test correction \cite{HolmBonferroni}, detailed in Section \ref{sec:multi-test}), $\gamma$ is defined as the probability of \textbf{not} having at least one point in the $\frac{\alpha}{n_{features}}$ upper tail of the distribution and the minimum number of samples $S$ to make is given by:
\begin{equation*}
    \label{eq:optimal-n-samples}
    \textit{S} = \ceil[\bigg]{\frac{\log\gamma}{log(1 - \frac{\alpha}{n_{features}})}}
\end{equation*}

@@

For each feature, we have now \textit{S} histograms, one for each sample. We compute the divergence between each of the \textit{S} histograms and the reference histogram. Figure \ref{fig:compute-sample-distances} illustrates this process as we apply our divergence or distance function for each sample's EMA-based histogram and the reference histogram, obtaining \textit{S} divergence measurements.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.4]{figures/compute-sample-distances.png}
      \caption[Compute sample's distance values]{Compute divergence values between reference and samples}
      \label{fig:compute-sample-distances}
    \end{center}
\end{figure}
These $S$ measurements are divergence measurements between random samples of data and the reference period. Hence, we claim we now know the distribution of expected distance values. Given a new distance value measured between another random histogram and the reference one, we can compute its probability and produce alerts if the probability is below a certain threshold. Note that the histograms for each sample are EMA-based histograms just like the streaming histogram. This ensures a certain degree of fidelity in the test we make in streaming because we measure the divergence between a random sample of data (our streaming histogram or any of the sample's histogram) and the reference one.

\subsubsection{Burn-In Period at System Initialization}
When we boot up our system in the streaming phase we initially have empty histogram aggregations. To avoid this burn-in period, we propose saving the last sample's histogram for each feature so that the streaming EMA-based histogram aggregation is initialized with its counts.

@@

\subsection{Streaming Phase} \label{sec:stream-phase}
In the streaming phase, the system will process event by event, update the streaming aggregations and periodically perform a change detection test.

For each feature, we initialize the approximate histogram with the last sample's data, from the batch phase. Then, for each incoming event and each of the event's features, we update our streaming histogram for that feature by adding the new value and then "cutting a slice" from the top of the histogram bars to simulate eviction, as described in Section \ref{sec:ema-hist}. This is all we do, regarding event processing.

We periodically perform a multi-feature change detection test. The test is done by first computing the divergence (here denoted as $div_x$) between the reference histogram for a feature \textit{x} and the current streaming histogram for $x$. This value is tested against the known distribution of distance values for feature \textit{x}. We test to see if the percentile of the computed distance relative to the distribution of divergence is above a certain percentile. The probability of a divergence value $div_x$ is:
\[pvalue(div_x) = 1 - percentile(div_x)\]
For example, if we define the minimum probability threshold to be 1\%, it means the system will raise an alert for distance values above the 99th-percentile or below a probability \textit{p-value} = 0.01. We alert a feature \textit{x} as divergent at a given timestamp if for a given probability threshold $\alpha$:
\[ percentile(div_x) > 1 - \alpha \]
or:
\begin{equation}
    \label{eq:alert-test}
    pvalue(div_x) < \alpha
\end{equation}

\subsubsection{Multiple Test Correction with Holm-Bonferroni} \label{sec:multi-test}
Note that we periodically apply this test (Equation \ref{eq:alert-test}) for each feature, effectively repeating the analysis. When considering several hypotheses, the problem of multiplicity arises \cite{MultiTestProblem-Pubmed, MultiTestProblem-Dickhaus2014, MultiTestProblem-Miller1966, multiple-test-correction-geoffrey}: the more hypotheses are checked, the higher the probability of obtaining false positives. In the literature, many multiple test correction methods are proposed, but we opted to use the Holm-Bonferroni \cite{HolmBonferroni} multiple test correction because it is a simple and easy to compute correction that can reduce the number of false negatives while having as target to control the number of false positives. Furthermore, it does not require to assume any independence between the tested hypothesis (as other methods do) which would almost surely be a wrong assumption for realistic datasets. In statistics, the Family-Wise Error Rate (FWER) \cite{MultivariateMT, MultitestTamhane2018AdvancesIP} is the probability of having one or more false positives when performing multiple hypotheses tests, \textit{i.e.}, considering one hypothesis true when it is not. The Holm-Bonferroni method is one of many approaches for controlling the Family-Wise Error Rate \textit{(FWER)} by adjusting the rejection criteria for each hypothesis (or test performed). The Holm-Bonferroni correction takes Equation \ref{eq:alert-test} and corrects the right-side threshold $\alpha$ for each feature. This correction is done by dividing the threshold $\alpha$ by a correction factor. First, we compute the \textit{p-value} of each feature. We then order our features by \textit{p-value}, in ascending order. Next, starting from the first feature, the one with the smallest \textit{p-value}, and proceeding until the end of the list of features, we check if:
\begin{equation}
    pvalue(val_x) < \alpha / (m - k)
    \label{eq:corrected-pvalue}
\end{equation}
with $\alpha$ as the probability to reject the null hypothesis, \textit{m} as the number of hypotheses to test (in our case the number of features) and \textit{k} as the 0-based index of the current feature. We stop at the first feature that fails this test, and every feature processed up until this one is considered to be a divergent feature and an alert per feature is logged with corresponding probability value.

@@

\subsection{Single-Feature Analysis}
We started by validating our system using single-feature datasets before moving to multi-feature scenarios to facilitate system validation, considering the complexities of multi-feature datasets and the problem of multiple test correction.

\subsubsection{Varying Sample Sizes}
We began with single-feature scenarios, experimenting with different tuple-based half-lives $n_{1/2}$. The half-life $n_{1/2}$ parameter in turn controls the size of the samples we make in the batch phase and the streaming EMA tail size. We set a half-life $n_{1/2}$ value and then multiply it by a constant factor of 4 to get the size of the samples to make. We process $4n_{1/2}$ events, meaning any further data we processed would make our oldest events have less than 6\% weight, following Algorithm \ref{pseudo:ema-hist}'s successively applied suppression factor. Processing $4n_{1/2}$ events means the oldest event's contribution is $({2^{- \frac{1}{n_{1/2}}}})^{4n_{1/2}} = {2^{- \frac{4n_{1/2}}{n_{1/2}}}} = 2^{-4} \approx 0.06 = 6\%$.

@@

Recall that from the 2.5 millionth event forward we change the distribution from a gaussian to an uniform one (Figure \ref{fig:timeseries-t1}). Figure \ref{fig:JSD-signal-01} shows the distance (JSD) values computed during streaming at each event of the 5 million events from dataset T1, using the outputs from the batch phase of experiment 01. We clearly see a change in JSD values, and we move from values close to 0 to values close to 1 (or even 1). Figure \ref{fig:JSD-signal-zoom-01} shows the same signal but for the first half of the streaming period (essentially a zoom-in). We see our JSD values before the change were very low (magnitudes of $10^{-2}$).
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-625.pdf}
      \caption{Whole time-series JSD signal and threshold}
      \label{fig:JSD-signal-01}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-zoom-625.pdf}
      \caption{Zoom-in on JSD signal and threshold}
     \label{fig:JSD-signal-zoom-01}
    \end{center}
\end{figure}
However, the change in JSD value is very abrupt as well. Almost instantly we go from close to 0 values to 1. We also had some false-positives, \textit{i.e.}, some JSD values above the threshold. This indicates to us that the window used is too small and sensitive to small changes, or that we can choose a more strict threshold.

In experiment 02, we use larger sampling windows. Remember our EMa tail size in streaming is proportional to the sampling windows. In the second experiment, we increase the size of the sampling windows by 100 times. Figure \ref{fig:JSD-signal-02} shows the JSD signal during streaming for dataset T1, using the outputs from the batch phase of experiment 02. Again, we see a change in signal from the middle of the streaming period forward. Figure \ref{fig:JSD-signal-zoom-02} shows the same signal for the first half of the streaming period.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-62500.pdf}
      \caption{Whole time-series JSD signal and threshold}
      \label{fig:JSD-signal-02}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-zoom-62500.pdf}
      \caption{Zoom-in on JSD signal and threshold}
     \label{fig:JSD-signal-zoom-02}
    \end{center}
\end{figure}
Notice that this time the growth of our signal is slower. Notice also that we do not have false positives in the first half of the signal. This is because we use a half-life, which means each event has a smaller contribution to the aggregation state or EMA tail. Hence, outliers will have less impact on the aggregation and corresponding signal value.

In experiment 03, we increase the half-life again, this time four times larger than in experiment 02. We would expect an even smoother growth of our signal and, again, no false positives. And indeed, judging by the corresponding signal computation and zoom-in, Figures \ref{fig:JSD-signal-03} and \ref{fig:JSD-signal-zoom-03}, respectively, the signal looks smoother and we see no false positives.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-250000.pdf}
      \caption{Whole time-series JSD signal and threshold}
      \label{fig:JSD-signal-03}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-zoom-250000.pdf}
      \caption{Zoom-in on JSD signal and threshold}
     \label{fig:JSD-signal-zoom-03}
    \end{center}
\end{figure}
However, the execution time was of course much larger. We will need to scale our experiments and do our batch analysis on more features and perform more samples, so we opt to use the 62,500 half-life from experiment 02 since it is a good compromise between batch execution time and EMA tail sensitivity to outlier values.

\subsubsection{Multiple Distribution Changes}
In this experiment, we wanted to see how our signal behaved when there were multiple change points, where each gaussian distribution was a scaled down version of the previous one. Figure \ref{fig:timeseries-r2} shows the time-series for the reference dataset R2 used in this experiment. It contained 5 million events with a single feature following a gaussian distribution with mean $\mu=100$ and standard deviation $\sigma=50$. The time-series for target dataset T2 is shown in Figure \ref{fig:timeseries-t2}. Notice that we change distributions at 2, 3 and 4 millionth events. Notice also that each subsequent distribution produces values whose domain is a subset of the previous distribution.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/timeseries-r2.pdf}
      \caption{Time-series for dataset R2}
      \label{fig:timeseries-r2}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/timeseries-t2.pdf}
      \caption{Time-series for dataset T2}
     \label{fig:timeseries-t2}
    \end{center}
\end{figure}
The target dataset T2 begins with the same distribution as the reference dataset R2, a gaussian distribution with mean $\mu=100$ and standard deviation $\sigma=50$. At the 2 millionth event mark, we use another Gaussian distribution, this time with mean $\mu=100$ and standard deviation $\sigma=30$. At the 3 millionth event, we change the standard deviation to be $\sigma=10$. From the 4 millionth event to the end, we use a standard deviation of $\sigma=5$. In this experiment, we used a half-life $n_{1/2}=62500$ and corresponding sample size of 250,000 events, like in the previous section's experiment 02. Figure \ref{fig:JSD-signal-test02} shows the computed JSD signal for the target dataset and Figure \ref{fig:JSD-signal-zoom-test02} zooms in on the first 2 millionth events, before any distribution is changed. We observe that once again we have no false positives in the first 2 million events, which is ideal. 
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-test02.pdf}
      \caption{Whole time-series JSD signal and threshold}
      \label{fig:JSD-signal-test02}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-zoom-test02.pdf}
      \caption{Zoom-in on first 2 million events}
     \label{fig:JSD-signal-zoom-test02}
    \end{center}
\end{figure}
Furthermore, we observe that the computed JSD signal varies accordingly with the changes in distribution. Despite us changing the distributions to subsets of the previous ones to try and deceive our method, our signal accurately represents these changes. In Figure \ref{fig:JSD-signal-test02}, we see several growth points that correspond to the points where the generating gaussian distribution parameters changed. Hence, we are able to detect changes in distribution parameter changes and quantify how big the change is (the closer to 1 our signal is).

\subsubsection{Multiple Distribution Changes, each a Superset of the Previous One}
In this experiment, we wanted to do the reverse of the previous one. In this experiment, there are multiple change points but each new distribution spans a domain that is a superset of the previous one. Figure \ref{fig:timeseries-r3} shows the reference period used, with 5 million events and one feature following a gaussian distribution of mean $\mu=100$ and standard deviation $\sigma=10$. Figure \ref{fig:timeseries-t3} shows the target period used, also with 5 million events and one feature. Like in the previous experiment, we change distributions at the 2, 3 and 4 millionth event mark. On the other hand, unlike the previous experiment, we now make subsequent distributions cover the previous ones. We change the reference gaussian used with mean $\mu=100$ and standard deviation $\sigma=10$ at the 2, 3 and 4 millionth mark, to use standard deviations $\sigma=30$, $\sigma=40$ and $\sigma=50$, respectively.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/timeseries-r3.pdf}
      \caption{Time-series for dataset R3}
      \label{fig:timeseries-r3}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/timeseries-t3.pdf}
      \caption{Time-series for dataset T3}
     \label{fig:timeseries-t3}
    \end{center}
\end{figure}
